\documentclass[11pt,a4paper]{report}

% Essential packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{setspace}

% Page geometry and spacing
\geometry{margin=1in}
\onehalfspacing

\begin{document}

% Chapter 1 content
\label{chap:1}
\chapter{Chapter 1: Introduction}

\section{Background and Motivation}

Stress is a common physiological and psychological response that significantly impacts human-computer interaction (HCI), health monitoring, and emotion recognition. Measuring a user's stress level reliably and unobtrusively is valuable in contexts like adaptive user interfaces and mental health assessment. Galvanic Skin Response (GSR), or electrodermal activity, is a recognized index of stress and arousal, reflecting sweat gland activity through skin conductance measurements \cite{Boucsein2012}. Traditional GSR monitoring techniques, however, rely on attaching electrodes to the skin (typically on the fingers or palm) to sense minute electrical conductance changes \cite{Fowles1981}. While effective in controlled settings, this contact-based approach has significant drawbacks: sensors can be intrusive and uncomfortable, often altering users' natural behaviour and emotional state \cite{Cacioppo2007}. In other words, the very act of measuring stress via contact sensors may itself induce stress or otherwise confound the measurements, raising concerns about ecological validity in HCI and ambulatory health scenarios \cite{Wilhelm2010}. Moreover, contact sensors tether participants to devices, limiting mobility and making longitudinal or real-world monitoring cumbersome.

These limitations motivate the pursuit of contactless stress measurement methods that can capture stress-related signals without any physical attachments, thereby preserving natural behaviour and comfort. Recent advances in sensing and computer vision suggest that it may be feasible to infer physiological stress responses using ordinary cameras and imaging devices, completely bypassing the need for electrode contact \cite{Picard2001}. Previous work in affective and physiological computing shows that visual cues—facial expressions, skin pallor, perspiration, and subtle movements—correlate with emotional arousal and stress levels \cite{Healey2005}. For example, thermal infrared imaging of the face can detect temperature changes linked to blood flow variations under stress, such as cooling of the nose tip due to vasoconstriction, in a non-contact manner. Likewise, high-resolution RGB video can monitor heart rate or breathing rate through subtle skin colour fluctuations and movements, as demonstrated in emerging remote photoplethysmography techniques \cite{Poh2010}.

\section{Research Problem and Objectives}

\subsection{Problem Context and Significance}

Recent advancements in contactless sensing pose a crucial question at the intersection of computer vision and psychophysiology: Can GSR-based stress measurements be approximated using only RGB camera video data? Specifically, can a simple video recording provide enough information to estimate physiological stress responses, eliminating the need for skin contact sensors? An affirmative answer could have significant implications. It would enable accessible stress monitoring through smartphones or laptop cameras, seamlessly integrating stress detection into everyday interactions and health monitoring, without the need for wearables or electrodes.

\subsection{Aim and Specific Objectives}

To investigate this research question, this project aims to develop and validate a system for contactless stress measurement. The primary objectives are to develop a multimodal data acquisition platform, utilise it to conduct a controlled stress-induction experiment, and analyse the resulting data to model GSR based on video features.

To achieve this, we first developed a multi-sensor data acquisition platform, named \textit{bucika\_gsr}. The system architecture spans two tightly integrated components: a custom Android mobile application and a desktop PC application. The Android app operates on a modern smartphone equipped with an attachable thermal camera module, simultaneously capturing both thermal and standard high-definition RGB video. Complementing this, the desktop PC application serves as the master controller, connecting via Bluetooth to a Shimmer3 GSR+ sensor to record the participant's ground-truth skin conductance in real-time.

A key objective was to ensure precise temporal alignment. The Android and PC components communicate over a wireless network, following a master--slave synchronisation protocol. The PC controller orchestrates the timing of recordings across all devices, achieving timestamp synchronisation on the order of a few milliseconds. Such tight synchronisation is crucial for our research, as it enables frame-by-frame correlation of physiological signals with visual cues captured on video \cite{Gravina2017}. The development of this platform involved several technical contributions, including a real-time multi-device synchronisation mechanism, an integrated framework for capturing heterogeneous data, and an extensible user interface architecture.

A further objective was to gather a high-quality dataset. Using the \textit{bucika\_gsr} platform, we conducted a controlled experiment where human participants underwent a standardised stress induction protocol (e.g., a time-pressured mental arithmetic task). Throughout each session, the system logged three synchronised data streams: continuous GSR signals, thermal video, and RGB video. The ultimate objective is to analyse this multimodal dataset. By examining the time-synchronised recordings, we can directly compare the GSR readings with visual data to determine what correlates of stress are present in the videos and quantify the limits of video-only stress assessment.

\section{Thesis Structure and Scope}

The document is organised as follows: Chapter~\ref{chap:2} reviews the multimodal physiological data collection platform, related work, the psychophysiology of stress responses, and recent advances in contactless physiological monitoring. Chapter~\ref{chap:3} defines the requirements and system analysis. Chapter~\ref{chap:4} details the design and implementation of the \textit{bucika\_gsr} platform. Chapter~\ref{chap:5} covers the evaluation and testing methodology and data analysis. Finally, Chapter~\ref{chap:6} concludes the thesis, discussing the findings concerning the research question, the limitations of the current approach, and potential directions for future research.

% Bibliography (minimal for standalone)
\bibliographystyle{ieeetr}
\bibliography{references}

\end{document}
